---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r echo = TRUE}
archive_name <- "activity.zip"
file_name <- "activity.csv"

if (!file.exists(file_name)) {
   unzip(zipfile = archive_name)
}

data <- read.csv(file = file_name)

data_steps <- data[!is.na(data$steps),]
data_steps$interval <- as.factor(data_steps$interval)

steps_per_day <- sapply(split(data_steps$steps, data_steps$date), sum)

average_steps <- sapply(split(data_steps$steps, data_steps$interval), mean)
```

## What is mean total number of steps taken per day?
```{r echo = TRUE}
hist(steps_per_day)
```

The mean value
```{r echo = TRUE}
mean(steps_per_day)
```
The median value
```{r echo = TRUE}
median(steps_per_day)
```

## What is the average daily activity pattern?
```{r echo = TRUE}
plot(x = names(average_steps), y = average_steps, type = "l")
```

The number of an interval having a maximum number of steps
```{r echo = TRUE}
names(average_steps)[max(average_steps)]
```

## Imputing missing values
```{r echo = TRUE}
sum(is.na(data$steps))
```
Creating of a copy data set
```{r echo = TRUE}
data1 <- data
```
Finding NA values in the new data set
```{r echo = TRUE}
to_fill <- is.na(data1$steps)
```
Filling NAs with the avarage numbers of steps for a given interval, although they can be absent.
```{r echo = TRUE}
data1[to_fill, ]$steps <- average_steps[data1[to_fill, ]$interval/5 + 1]
```
Finding values that are still NA
```{r echo = TRUE}
to_fill2 <- is.na(data1$steps)
```
Filling NAs with average values of steps per day, that are existing always.
```{r echo = TRUE}
data1[to_fill2, ]$steps <- steps_per_day[data1[to_fill2, ]$date]/17568

steps_per_day1 <- sapply(split(data1$steps, data1$date), sum)

hist(steps_per_day1)
```

The mean value
```{r echo = TRUE}
mean(steps_per_day1)
```
The median value
```{r echo = TRUE}
median(steps_per_day1)
```
There are more small numbers added for most of NAs where at night when small numbers od steps are done. The mean value has changed a little, the median value is the same.

## Are there differences in activity patterns between weekdays and weekends?
Changing the locale to english one to havs proper day names. Then creating a new column in the data set and filling it with the factor (weekday/weekend).
```{r echo = TRUE}
Sys.setlocale("LC_TIME", "English")
data1$daytype <- weekdays(as.Date(as.character(data1$date)))
data1$daytype[data1$daytype %in% c("Saturday", "Sunday")] <- "weekend"
data1$daytype[!data1$daytype %in% c("weekend")] <- "weekday"
data1$daytype <- as.factor(data1$daytype)
library(ggplot2)
qplot(interval, steps, data = data1, facets = daytype~., stat="summary", fun.y = "mean", geom = "line")
```